\TITLE

### 2021-01-12 vonLaszewski-covid.tex
Updates from Overleaf

% spacing
% outcomment this for production
\usepackage{setspace}
\usepackage{lineno}
\linenumbers
\doublespace
% above needs to be commented out for production

@@

\begin{table}[!hptb]
\caption{Covariate Risk factor abbreviations as used in this study.}
\label{tab:risk-factors}
\bigskip
\resizebox{0.90\textwidth}{!}{
\centering
\begin{tabular}{ll}
\toprule
Risk Factor & Description \\
\midrule
 NONE & No risk factor used\\
 ARTHRITIS 	  & Percent reported with arthritis \\
 BINGE 		  & Percent reported with binge drinking \\
 BPHIGH 	  & Percent reported with  high blood pressure\\
 BPMED 		  & Percent reported taking blood pressure medication\\
 CANCER 	  & Percent reported as cancer patients\\
 CASTHMA 	  & Percent reported with current asthma\\
 CHD 		  & Percent reported with coronary heart disease\\
 CHECKUP 	  & Percent reported with health checkup\\
 CHOLSCREEN   & Percent reported with cholesterol screening\\
 COPD 		  & Percent reported with chronic obstructive pulmonary disease\\
 CSMOKING 	  & Percent reported as currently smoking\\
 DIABETES 	  & Percent reported with diabetes\\
 ESTBEDS 	  & Number of estimated beds in hospitals\\
 HIGHCHOL 	  & Percent reported with high cholesterol\\
 INSURANCE 	  & Percent reported with insurance\\
 KIDNEY 	  & Percent reported with kidney disease\\
 LPA 		  & Percent reported with no leisure-time physical activity \\
 MHLTH 		  & Percent reported with not good mental health\\
 NBEDS 		  & Number of beds in hospitals \\
 NBEDS/1000   & Number of beds per 1000 people\\
 NHOSP 		  & Number of hospitals\\
 OBESITY 	  & Percent reported with  population that is obese\\
 PHLTH        & Percent reported with physical health issues\\
 PVI 		  & Pandemic vulnerability index \\
 STROKE 	  & Percent reported with stroke\\
 BLACK	  & Percent of Blacks in the population\\
 POVERTY & Percent of poverty in the population\\
 SENIOR  & Percent of seniors in the population\\
 SVI\_MINORITY 	  & Percent of social vulnerability index in the minority population\\
 SVI\_OVERALL 	  & Percent of social vulnerability index in the overall population\\
\bottomrule
\end{tabular}
}
\end{table}

@@

    \begin{subfigure}{.49\textwidth}
      \centering
      \includegraphics[width=1.0\linewidth]{images/predict/Fig4a.pdf}  
      \caption{Prediction acuracy for deaths.}
      \label{fig:magic-a}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
      \centering
      \includegraphics[width=1.0\linewidth]{images/predict/Fig4b.pdf}        
      \caption{Prediction acuracy for cases.}
      \label{fig:magic-b}
    \end{subfigure}

@@

    \includegraphics[width=1.0\textwidth]{images/boxwhisker/boxplot_deathNEW2.pdf}
    \vspace{-1cm}
    \caption{Influence of risk factors on the accuracy of the
      prediction of COVID-19 death. The y-axis represents the
      cumulative error over all input data for the cities.}
    \label{fig:box-death}
\end{figure}

@@

\item[Order for cases sorted by minimum model error:] NONE,
  POP\_DENSITY\_2010, PHLTH, INSURANCE, BLACK,
  HISPANICS, NHOSP, CHOLSCREEN, DIABETES, STROKE,
  CHD, CHECKUP, NBEDS, SVI\_OVERALL, KIDNEY, CASTHMA, BPMED,
  LPA, CSMOKING, ARTHRITIS, POVERTY, ESTBEDS, MHLTH, SENIOR, COPD,
  CANCER, NBEDS/1000, BPHIGH, HIGHCHOL, BINGE, OBESITY, SVI\_MINORITY

\item[Order for deaths sorted by minimum model error:] NONE,
  SVI\_OVERALL, CSMOKING, HISPANICS, POP\_DENSITY\_2010,
  SENIOR, CHOLSCREEN, CASTHMA, INSURANCE, MHLTH, LPA,
  POVERTY, OBESITY, CANCER, BLACK, NBEDS/1000, DIABETES, ARTHRITIS,
  NBEDS, NHOSP, BPHIGH, CHECKUP, PHLTH, CHD, STROKE, HIGHCHOL,
  ESTBEDS, KIDNEY, BINGE, BPMED, COPD, SVI\_MINORITY
\end{description}

@@

We observed a different scenario for deaths, in which we only
identified two risk factors, namely the Social Vulnerability Index
(SVI) and chain-smoking, that, when integrated into our deep learning
model, leads to better prediction results. However, the accuracy of
the overall prediction of deaths is far less precise than that of the
cases. To showcase this, we have included Figures
\ref{fig:error-case-popdensity} and \ref{fig:error-death-popdensity}
that show the respective errors.

@@

This can be explained by our input data distribution between cases and deaths in which we noted higher
fluctuations relative to the overall value in cases of deaths. 
A mitigation to this issue
would be using a seven day average over the analyzed period. However,
we have not done this on purpose for this analysis as we wanted to
identify how a covariate enhanced LSTM behaves that includes risk
factors based on the daily fluctuations. This was done to avoid and
showcase any needed preproccessing on the data and to identify the
capability of the deep learning framework while adapting to the
fluctuations without any special activities. Previously, we have
already shown in Section \ref{sec:emperical} that for smooth data
inputs the prediction is very accurate.  Using fluctuation data allows
our framework for an automated ingest of data on a daily basis to
enable data fusion of new cases and death information. Hence, we can
re-train the model with new incoming data every day. Through repeated
training experiments, we found the LSTM hyperparameters such as epochs
and dropout values that work well for our experiments.

@@

Notably, we identified that a jump occurs when we sort the top 10
model predictions at around 330 models from a total of 3400
automatically generated models. This also underscores the need to run
our deep learning model multiple times to obtain best parameter
candidates.  Furthermore, we see that the cumulative error for cases
is about half the size as for death. This is caused by cases having a
larger number of samples than those from deaths.

@@

%    \begin{minipage}{.45\textwidth}
%        \centering
%        \includegraphics[width=1.0\textwidth]{images/predict/PlaceTop5_Cases.pdf}
%        \vspace{-1cm}
%        \caption{Top 5 predictions from all cases over all risk factors.}
%        \label{fig:place-top5-cases}
%    \end{minipage}
%    \ \
%    \begin{minipage}{.45\textwidth}
%        \centering
%        \includegraphics[width=1.0\textwidth]{images/predict/PlaceTop5_Death.pdf}
%        \vspace{-1cm}
%        \caption{Top 5 predictions from all deaths  over all risk factors.}
%        \label{fig:place-top5-death}
%    \end{minipage}
    \begin{subfigure}{.95\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{images/predict/PlaceTop10_DeathNEW3.pdf}
        \caption{Cumulative error for deaths.}
        \label{fig:place-top10-death}
    \end{subfigure}
    \begin{subfigure}{.95\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{images/predict/PlaceTop10_CasesNEW3.pdf}
        \caption{Cumulative error for cases}
        \label{fig:place-top10-cases}
    \end{subfigure}

@@

The comprehensive analysis showcases the feasibility of our
approach. Based on the outcomes reported it can lead to an improved
prediction once we integrate risk factors in addition to the
time-dependent data such as cases and death resulting from
COVID-19. While the improvements observed were modest, they do present
a concrete way of improving the forecast models. Inclusion of further
putative factors from the ongoing worldwide studies on Covid-19 will
only strengthen the future applications of our integrative framework.

>>>>

% spacing
% outcomment this for production
\usepackage{setspace}
\usepackage{lineno}
\linenumbers
%\doublespace
% above needs to be commented out for production

@@

\begin{table}[!hptb]
\caption{Covariate Risk factor abbreviations as used in this study.}
\label{tab:risk-factors}
\bigskip
\resizebox{0.90\textwidth}{!}{
\centering
\begin{tabular}{ll}
\toprule
Risk Factor & Description \\
\midrule
 ARTHRITIS 	  & Percent reported with arthritis \\
 BINGE 		  & Percent reported with binge drinking \\
 BLACK	  & Percent of Blacks in the population\\
 BPHIGH 	  & Percent reported with  high blood pressure\\
 BPMED 		  & Percent reported taking blood pressure medication\\
 CANCER 	  & Percent reported as cancer patients\\
 CASTHMA 	  & Percent reported with current asthma\\
 CHD 		  & Percent reported with coronary heart disease\\
 CHECKUP 	  & Percent reported with health checkup\\
 CHOLSCREEN   & Percent reported with cholesterol screening\\
 COPD 		  & Percent reported with chronic obstructive pulmonary disease\\
 CSMOKING 	  & Percent reported as currently smoking\\
 DIABETES 	  & Percent reported with diabetes\\
 ESTBEDS 	  & Number of estimated beds in hospitals\\
 HIGHCHOL 	  & Percent reported with high cholesterol\\
 HISPANIC	  & Percent of Hispanic in the population\\
 INSURANCE 	  & Percent reported with insurance\\
 KIDNEY 	  & Percent reported with kidney disease\\
 LPA 		  & Percent reported with no leisure-time physical activity \\
 MHLTH 		  & Percent reported with not good mental health\\
 NBEDS 		  & Number of beds in hospitals \\
 NBEDS/1000   & Number of beds per 1000 people\\
 NHOSP 		  & Number of hospitals\\
 NONE & No risk factor used\\
 OBESITY 	  & Percent reported with  population that is obese\\
 PHLTH        & Percent reported with physical health issues\\
 POP\_DENSITY\_2010 & Population Density from the 2010 Census\\
 POVERTY & Percent of poverty in the population\\
 % PVI 		  & Pandemic vulnerability index \\
 SENIOR  & Percent of seniors in the population\\
 STROKE 	  & Percent reported with stroke\\
 SVI\_MINORITY 	  & Percent of social vulnerability index in the minority population\\
 SVI\_OVERALL 	  & Percent of social vulnerability index in the overall population\\
\bottomrule
\end{tabular}
}
\end{table}

@@

    \begin{subfigure}{.49\textwidth}
      \centering
      \includegraphics[width=1.0\linewidth]{images/predict/Fig4a.pdf}  
      \caption{Prediction accuracy for deaths.}
      \label{fig:magic-a}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
      \centering
      \includegraphics[width=1.0\linewidth]{images/predict/Fig4b.pdf}        
      \caption{Prediction accuracy for cases.}
      \label{fig:magic-b}
    \end{subfigure}

@@

    \includegraphics[width=1.0\textwidth]{images/boxwhisker/boxplot_deathNEW2.pdf}
    \vspace{-1cm}
    \caption{Influence of risk factors on the accuracy of the
      prediction of COVID-19 deaths. The y-axis represents the
      cumulative error over all input data for the cities.}
    \label{fig:box-death}
\end{figure}

@@

\item[Order for cases sorted by minimum model error:] 
  POP\_DENSITY\_2010, PHLTH, BLACK, HISPANIC, NHOSP, INSURANCE, 
  CHOLSCREEN, DIABETES, STROKE, CHD, CHECKUP, NBEDS, SVI\_OVERALL, 
  KIDNEY, CASTHMA, BPMED, LPA, CSMOKING, ARTHRITIS, POVERTY, ESTBEDS, 
  MHLTH, COPD, SENIOR, CANCER, NBEDS/1000, BPHIGH, NONE, HIGHCHOL, 
  BINGE, OBESITY, SVI\_MINORITY

\item[Order for deaths sorted by minimum model error:] 
  SVI\_OVERALL, CSMOKING, HISPANIC, NONE, POP\_DENSITY\_2010, 
  SENIOR, CHOLSCREEN, CASTHMA, INSURANCE, MHLTH, LPA, 
  POVERTY, OBESITY, CANCER, NBEDS/1000, BLACK, DIABETES, ARTHRITIS, 
  NBEDS, NHOSP, BPHIGH, PHLTH, CHECKUP, CHD, STROKE, HIGHCHOL, 
  ESTBEDS, KIDNEY, BINGE, BPMED, COPD, SVI\_MINORITY

\end{description}

@@

We observed a different scenario for deaths, in which we only
identified three risk factors, namely the Social Vulnerability Index
(SVI), chain-smoking, and Hispanic that, when integrated into our deep learning
model, leads to better prediction results. However, the accuracy of
the overall prediction of deaths is far less precise than that of the
cases. 
To showcase this, we have included Figures
\ref{fig:error-case-popdensity} and \ref{fig:error-death-popdensity}
that show the respective errors for example for the population density.

@@

This can be explained by our input data distribution between cases and deaths in which we noted higher
fluctuations relative to the overall value in cases of deaths. 
A mitigation to this issue
would be using a seven day average over the analyzed period. However,
we have not done this on purpose for this analysis as we wanted to
identify how a covariate enhanced LSTM behaves that includes risk
factors based on the daily fluctuations. This was done to avoid and
showcase any needed preproccessing on the data and to identify the
capability of the deep learning framework while adapting to the
fluctuations without any special activities. Previously, we have
already shown in Section \ref{sec:emperical} that for smooth data
inputs the prediction is very accurate.  Using fluctuation data allows
our framework for an automated ingest of data on a daily basis to
enable data fusion of new cases and deaths information. Hence, we can
re-train the model with new incoming data every day. Through repeated
training experiments, we found the LSTM hyperparameters such as epochs
and dropout values that work well for our experiments.

@@

Notably, we identified that a jump occurs when we sort the top 10
model predictions at around 320 models from a total of 3200
automatically generated models. This also underscores the need to run
our deep learning model multiple times to obtain best parameter
candidates.  Furthermore, we see that the cumulative error for cases
is about half the size as for deaths. This is caused by cases having a
larger number of samples than those from deaths.

@@

%    \begin{minipage}{.45\textwidth}
%        \centering
%        \includegraphics[width=1.0\textwidth]{images/predict/PlaceTop5_Cases.pdf}
%        \vspace{-1cm}
%        \caption{Top 5 predictions from all cases over all risk factors.}
%        \label{fig:place-top5-cases}
%    \end{minipage}
%    \ \
%    \begin{minipage}{.45\textwidth}
%        \centering
%        \includegraphics[width=1.0\textwidth]{images/predict/PlaceTop5_Death.pdf}
%        \vspace{-1cm}
%        \caption{Top 5 predictions from all deaths  over all risk factors.}
%        \label{fig:place-top5-death}
%    \end{minipage}
    \begin{subfigure}{.95\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{images/predict/PlaceTop10_DeathNEW3.pdf}
        \caption{Cumulative error for deaths.}
        \label{fig:place-top10-death}
    \end{subfigure}
    \begin{subfigure}{.95\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{images/predict/PlaceTop10_CasesNEW3.pdf}
        \caption{Cumulative error for cases.}
        \label{fig:place-top10-cases}
    \end{subfigure}

@@

The comprehensive analysis showcases the feasibility of our
approach. Based on the outcomes reported it can lead to an improved
prediction once we integrate risk factors in addition to the
time-dependent data such as cases and deaths resulting from
COVID-19. While the improvements observed were modest, they do present
a concrete way of improving the forecast models. Inclusion of further
putative factors from the ongoing worldwide studies on Covid-19 will
only strengthen the future applications of our integrative framework.
